---
title: "5206.hw7.cm3700"
author: "Chi.Ma,cm3700"
date: "11/27/2017"
output: html_document
---

```{r}
setwd("~/Desktop/5206/HW")
df<-read.csv("moretti.csv")
```
#Part1
```{r}
poisLoglik<-function(data,lambda){
  return(sum(dpois(x=data,lambda=lambda,log = TRUE)))
}
poisLoglik(c(1,0,0,1,1),1)

```
#part2
Write a function count new genres which takes in a year, and returns the number of new genres which appeared in that year: 0 if there were no new genres that year, 1 if there was one, 3 if there were three, etc. What should the values be for 1803 and 1850?
```{r}
count_new_genres<-function(year){
  return(sum(df$Begin==year))
}
count_new_genres(1803)
count_new_genres(1850)

```
#part3
Create a vector, new genres, which counts the number of new genres which appeared in each year of the data, from 1740 to 1900. What positions in the vector correspond to the years 1803 and 1850? What should those values be? Is that what your vector new   genres has for those years?
```{r}

years<-seq(1740,1900)
new_genres<-rep(NA,length(years))

for(i in 1:length(years)){
  new_genres[i]<-count_new_genres(years[i])
}

which(years==1803)
which(years==1850)
new_genres[years==1805]
new_genres[years==1850]

```

```{r}
#The positions for year 1803 and 1850 are 64 and 111, and the values are 0 and 3 respectively.
```
#Part4
Plot poisLoglik as a function of λ on the new_genres data. (If the maximum is not at λ = 0.273, you’re doing something wrong.)
```{r}
lambda<-seq(0,3,by=0.001)
likelihood<-c()
for(i in 1:length(lambda)){
  likelihood<-c(likelihood,poisLoglik(new_genres,lambda[i]))
}
plot(lambda,likelihood,type = "l",,xlab = "Lambda",ylab = "Log Likelihood",main = "Log Likelihood/Lambda")

lambda[which.max(likelihood)]

```
#Part5
Use nlm() to maximize the log likelihood to check the λ = 0.273 value suggested in the previous question. Hint: you may need to rewrite your function from (i.) with some slight alterations

```{r}
poisLoglik_2<-function(lambda,data){
  return(-sum(dpois(data,lambda=lambda,log = TRUE)))
}
nlm(poisLoglik_2, 1, new_genres)
```
#Part6
To investigate whether genres appear in bunches or randomly, we look at the spacing between genre births. Create a vector, intergenre intervals, which shows how many years elapsed between new genres appearing. (If two genres appear in the same year, there should be a 0 in your vector, if three genres appear in the same year your vector should have two zeros, and so on. For example if the years that new genres appear are 1835,1837,1838,1838,1838 your vector should be 2,1,0,0.) What is the mean of the time intervals between genre appearances? The standard deviation? The ratio of the standard deviation to the mean, called the coefficient of variation? Hint: The diff() function might help you here. Check out ?diff.

```{r}
intergenre_intervals<-diff(df$Begin)
mu<-mean(intergenre_intervals)
sd<-sd(intergenre_intervals)
cv<-(sd/mu)
mu
sd
cv
```
```{r}
#mean = 3.44186, sd = 3.705224, coefficient of variance = 1.076518
```
#Part7
For a Poisson process, the coefficient of variation is expected to be around 1. However, that calculation doesn’t account for the way Moretti’s dates are rounded to the nearest year, or tell us how much the coefficient of variation might fluctuate. We will handle both of these by simulation.
a.Write a function which takes a vector of numbers, representing how many new genres appear in each year, and returns the vector of the intervals between appearances. Check that your function works by seeing that when it is given new genres, it returns intergenre intervals.
```{r}
intergenre_intervals_cals<-function(data){
  names(data)<-1:length(data)
  data<-data[data!=0]
  years<-as.numeric(rep(names(data),data))
  data_output<-diff(years)
  return(data_output)
}
all(intergenre_intervals_cals(new_genres)==intergenre_intervals)
```

b.Write a function to simulate a Poisson process and calculate the coefficient of vari- ation of its inter-appearance intervals. It should take as arguments the number of years to simulate and the mean number of genres per year. It should return a list, one component of which is the vector of inter-appearance intervals, and the other their coefficient of variation. Run it with 161 years and a mean of 0.273; the mean of the intervals should generally be between 3 and 4.
```{r}
CV<-function(years,mu){
  samples<-rpois(years,mu)
  sample.interval<-intergenre_intervals_cals(samples)
  cv.interval<-sd(sample.interval)/mean(sample.interval)
  sample.output<-list(sample.interval=sample.interval,cv.interval=cv.interval)
  return(sample.output)
}
CV(161,0.273)

```
#Part8
Run your simulation 10, 000 times, taking the coefficient of variation (only) from each. (This should take less than two minutes to run.) What fraction of simulations runs have a higher coefficient of variation than Moretti’s data?

```{r}
CV10K<-rep(NA,10000)
for(i in 1:10000){
  CV10K[i]<-CV(161,0.273)$cv.interval
}
mean(CV10K>cv)
```
#Part9
Explain what this does and does not tell you about the conjecture that genres tend to appear together in burst?

```{r}
#The result tells us the the new genres did not appear together in bursts and if new genres did appear in a possion distribution, 23.42% of chance we would see that results are more clustered than the Moretti's data.
```
