---
title: "Lab 5"
author: "Chi Ma, CM3700"
date: "November 10, 2017"
output:
  html_document:
    df_print: paged
---

# Instructions 
Make sure that you upload an RMarkdown file to the canvas page (this should have a .Rmd extension) as well as the PDF output after you have knitted the file (this will have a .pdf extension). The files you upload to the Canvas page should be updated with commands you provide to answer each of the questions below.  You can edit this file directly to produce your final solutions.  The lab is due 11:59pm on Saturday, November 9th.    

# Goal

The goal of this lab is to investigate the empirical behavior of a common hypothesis testing procedure through simulation using R. We consider the traditional two-sample t-test.  

# Two-Sample T-Test


Consider an experiment testing if a 35 year old male's heart rate statistically differs between a control group and a dosage group. Let $X$ denote the control group and let $Y$ denote the drug group.  One common method used to solve this problem is the two-sample t-test.  The null hypothesis for this study is:
$$H_0:\mu_1-\mu_2=\Delta_0,$$


# Tasks

1) Using the **R** function **t.test**, run the two sample t-test on the following simulated dataset.  Note that the **t.test** function defaults a two-tailed alternative.  Also briefly interpret the output.    
```{r}
rm(list = ls())
set.seed(5)
sigma=5
Control <- rnorm(30,mean=10,sd=sigma)
Dosage <- rnorm(35,mean=12,sd=sigma)
t.test(x=Control, y=Dosage)

```
```{r}
#t-value = -1.9684, p-value = 0.05349
#Failed to reject null hypothesis since p-value is larger than 5%. 
```
2)  Write a function called **emperical.size** that simulates **R** different samples of $X$ for control and **R** different samples of $Y$ for the drug group and computes the proportion of test statistics that fall in the rejection region.  The function should include the following:  

I started the function below: 
```{r}
# X contrl group Y drug group
emperical.size <- function(R=10000,
                           mu1=0,mu2=0,
                           sigma1=1,sigma2=1,
                           m=30,n=30,
                           level=.05,
                           value=0,
                           direction="Two") {
  
  #Define empty lists
  statistic.list <- rep(0,R)
  pvalue.list <- rep(0,R)
  
    for (i in 1:R) {
    
    #Sample realized data 
    Control <- rnorm(m,mean=mu1,sd=sigma1)
    Dosage <- rnorm(n,mean=mu2,sd=sigma2)
    
    #Testing values
    testing.procedure <- t.test(Control,Dosage,mu=value)
    statistic.list[i] <-testing.procedure$statistic 
    pvalue.list[i] <- testing.procedure$p.value
   
  
   }
   size <- mean(pvalue.list<=level)
     
  return(list(statistic.list=statistic.list,pvalue.list=pvalue.list,emperical.size=size)) 
  
}
```
Evaluate your function with the following inputs 
**R=10**,**mu1=10**,**mu1=12**,**sigma1=5** and **sigma2=5**. 
```{r}
emperical.size(R=10,mu1=10,mu2=12,sigma1=5,sigma2=5)

```



3) Assuming the null hypothesis $$H_0:\mu_1-\mu_2=0$$ is true, compute the empirical size using 10,000 simulated data sets.  Use the function **emperical.size** to accomplish this task and store the object as **sim**.  Output the empirical size quantity **sim$size**.  Comment on this value.  What is it close to?     

  **Note:**  use **mu1=mu1=10** (i.e., the null is true).  Also set **sigma1=5**,**sigma2=5** and **n=m=30**. 
```{r}
sim<-emperical.size(R=10000,mu1=10,mu2=10,sigma1 = 5,sigma2 = 5)
sim$emperical.size
```

4) Plot a histogram of the simulated P-values, i.e., **hist(sim$pvalue.list)**.  What is the probability distribution shown from this histogram?  Does this surprise you? 
```{r}
hist(sim$pvalue.list)
```
```{r}
#The probability distribution is close to uniform distribution, this histogram does suprize me because I didn't expect the p-values are distributed so evenly.
```
5) Plot a histogram illustrating the empirical sampling sampling of the t-statistic, i.e., **hist(sim$statistic.list,probability =TRUE)**.  What is the probability distribution shown from this histogram?    
```{r}
hist(sim$statistic.list,probability = TRUE)
```
```{r}
#The histogram is close to a normal distribution.
```
6) Run the following four lines of code:

      **emperical.size(R=1000,mu1=10,mu1=10,sigma1=5,sigma2=5)$emperical.size**

      **emperical.size(R=1000,mu1=10,mu1=12,sigma1=5,sigma2=5)$emperical.size**

      **emperical.size(R=1000,mu1=10,mu1=14,sigma1=5,sigma2=5)$emperical.size**

      **emperical.size(R=1000,mu1=10,mu1=16,sigma1=5,sigma2=5)$emperical.size**
      
      Comment on the results.  
```{r}
emperical.size(R=1000,mu1=10,mu2 =10,sigma1=5,sigma2=5)$emperical.size

emperical.size(R=1000,mu1=10,mu2=12,sigma1=5,sigma2=5)$emperical.size

emperical.size(R=1000,mu1=10,mu2=14,sigma1=5,sigma2=5)$emperical.size

emperical.size(R=1000,mu1=10,mu2=16,sigma1=5,sigma2=5)$emperical.size
```
```{r}
#When the difference between mu1 and mu2 increases, emperical.size increases and more test statistics are fall into rejection region.
```
7) Run the following four lines of code:

      **emperical.size(R=10000,mu1=10,mu2=12,sigma1=10,sigma2=10,m=10,n=10)$emperical.size**

      **emperical.size(R=10000,mu1=10,mu2=12,sigma1=10,sigma2=10,m=30,n=30)$emperical.size**

      **emperical.size(R=10000,mu1=10,mu2=12,sigma1=10,sigma2=10,m=50,n=50)$emperical.size**

      **emperical.size(R=10000,mu1=10,mu2=12,sigma1=10,sigma2=10,m=100,n=100)$emperical.size**
      
      Comment on the results.  
```{r}
emperical.size(R=10000,mu1=10,mu2=12,sigma1=10,sigma2=10,m=10,n=10)$emperical.size
emperical.size(R=10000,mu1=10,mu2=12,sigma1=10,sigma2=10,m=30,n=30)$emperical.siz
emperical.size(R=10000,mu1=10,mu2=12,sigma1=10,sigma2=10,m=50,n=50)$emperical.size
emperical.size(R=10000,mu1=10,mu2=12,sigma1=10,sigma2=10,m=100,n=100)$emperical.size
```      
```{r}
# When the sample size increases for both group, emperical size also increases. Alothough this effect is not as strong as the changes in mu, but this also shows that more test statistics are fall into rejecton region. 

```











